Estimating demographic bias on tests of children's early vocabulary

### 2022-09-14 paper/manuscript.Rmd
credit and abstract edits

author: 
  - name          : "George Kachergis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Jane Stanford Way"
    email         : "gkacherg@stanford.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Nathan Francis"
    affiliation   : "1"
    role:
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Michael C. Frank"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Supervision"

@@

abstract: |
  Children's early language skill has been linked to later educational outcomes, making
  it important to measure early language accurately. Parent-reported instruments such as the 
  Communicative Development Inventories (CDIs) have been shown to provide reliable and valid
  measures of children's aggregate early language skill. However, CDIs contain
  hundreds of vocabulary items, some of which may not be heard (and thus learned) equally
  often by children of varying backgrounds. This study used a database of American English 
  CDIs to identify words demonstrating strong bias for particular demographic groups of children, on 
  dimensions of sex (male vs. female), race (white vs. non-white), and maternal education 
  (high vs. low). For each dimension, we identified dozens of strongly biased items, and showed that 
  eliminating even some of these items can reduce the magnitude of differences between groups. 
  Additionally, we investigated how well the relative frequency of words spoken to young girls vs.
  boys predicted sex-based word learning bias, and discuss possible sources of demographic
  differences in early word learning.

>>>>

author: 
  - name          : "George Kachergis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Jane Stanford Way"
    email         : "gkacherg@stanford.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Methodology"
      - "Software"
      - "Investigation"
      - "Data Curation"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Visualization"
      - "Supervision"
  - name          : "Nathan Francis"
    affiliation   : "1"
    role:
      - "Software"
      - "Investigation"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Michael C. Frank"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Methodology"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Supervision"
      - "Funding acquisition"

@@

abstract: |
  Children's early language skill has been linked to later educational outcomes, making
  it important to measure early language accurately. Parent-reported instruments such as the 
  Communicative Development Inventories (CDIs) have been shown to provide reliable and valid
  measures of children's aggregate early language skill. However, CDIs contain
  hundreds of vocabulary items, some of which may not be heard (and thus learned) equally
  often by children of varying backgrounds. This study used a database of American English 
  CDIs to identify words demonstrating strong bias for particular demographic groups of children, on 
  dimensions of sex (male vs. female), race (white vs. non-white), and maternal education 
  (high vs. low). For each dimension, many items showed bias; removing these items slightly
  reduced the magnitude of group differences, but did not eliminate them. 
  Additionally, we investigated how well the relative frequency of words spoken to young girls vs.
  boys predicted sex-based word learning bias, and discuss possible sources of demographic
  differences in early word learning.
